### 2017-01-16 doc/papers/summary/HFStability-summary.tex

\documentclass{revtex4}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{graphicx}

\begin{document}
\title{Summary Hartree-Fock Stability of HEG}
\author{Evan Curtin}
\maketitle

\section{Background}
\cite{Guiliani2005}

\section{References}
\bibliography{../heg_references}

\end{document}